Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-13 bem.tex
added to intro

\begin{figure*}
  \centering
  \input{img/triangles}
  \caption{The four types of intersection of two triangles: the triangles can be (L to R) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle. In the first case, standard quadrature is used. In the other three cases, singular quadrature rules must be used.}
  \label{fig:triangles}
\end{figure*}

>>>>

\begin{figure*}
  \centering
  \input{img/triangles}
  \caption{The four types of intersection of two triangles: the triangles can be (left to right) not neighbours, neighbours adjacent via a vertex, neighbours adjacent via an edge, or the same triangle. In the first case, standard quadrature is used. In the other three cases, singular quadrature rules must be used.}
  \label{fig:triangles}
\end{figure*}

### 2021-04-13 intro.tex
added to intro

>>>>

\subsection{The three language problem}
The original BEM++ was troubled by what is often called the two language problem. It is common for programming language to be either easy for humans to write (e.g., Python) or easy for computers to run and acheive high performance (e.g., C++). It is not common, however, for a language to do both of these. Due to this, it is common in scientific computing libraries to write a library in a fast low-level language such as C++, while providing a user interface in a higher-level language such as Python.

This was the model followed by BEM++, but the problem would perhaps better be described as a three language problem: as well as the Python and C++ code contained in libraries like this, it is common to also include a significant amount of code in a third interfacing language. In the case of BEM++, this third language was Cython, an extension of Python with C data types that can be compiled to include functionality from C++ libraries.

Due to the three languages involved, making changes to the library would often mean having to duplicate changed in three places, with many class structures duplicated in all three languages. This made seemingly simple changes in onorous tasks, and provided a barrier to new members of the community looking to contribute to the open-source project.

\subsection{Delegating computations with PyOpenCL}
Prompted by a desire to simplify the library as well as to be able to run on a wide range of CPU and GPU devices, we began a full re-write in 2018, which led to Bempp-cl. The aims of this rewrite were to support explicit single-instruction-multiple-data (SIMD) optimization on CPUs with various instruction lengths, be able to offload computations to AMD, Intel, and Nvidia GPUs, and to base the complete codebase on Python. These aims naturally led to the choice of building a Python library based around OpenCL (using the PyOpenCL interface) and Numba.

In addition to the performance benefits, the library redesign has greatly improved the issues related to the three language problem. Both OpenCL and Numba are use to compile functions. Each function is provided all the data is needs as inputs, so there is no need to duplicate any class structure outside Python. The need for a third interfacing language is removed, as the interfacing to the OpenCL kernels is handled by PyOpenCL.

Boundary element methods are particularly suited to this library model, as the main performance critical task is the computation of discrete operators. Once fast kernels for this have been implemented, the remaining functionality of the library can be written entirely in Python without any great detriment to performance.
